Do we need detailed demographic data to forecast plant population responses to climate change?

### 2016-06-06 manuscript/supporting_information.Rmd
updates some text

\begin{align}
&\mathtt{data} \qquad &y_{i,q,t}^{S} \sim \text{Bernoulli}(s_{i,q,t}) \\
&\mathtt{process} \qquad &\text{logit}(s_{i,q,t}) = \beta_{0,t} + \beta_{s,t}x_{i,q,t} + \beta_Q + \textbf{z}'_t \boldsymbol{\beta}_c + \beta_{d,1} w_{i,t} + \beta_{d,2} (x_{i,q,t}w_{i,q,t}) \\
&\mathtt{parameters} \qquad &\beta_{0,t} \sim \text{Normal}(\tilde{\beta_0}, \sigma_{\beta_0}^2 \textbf{I})\\
& \qquad & \beta_{s,t} \sim \text{Normal}(\tilde{\beta_s}, \sigma_{\beta_s}^2 \textbf{I}) \\
& \qquad & \boldsymbol{\beta_{Q}} \sim \text{Normal}(0, \sigma_Q^2 \textbf{I}) \\
& \qquad & \boldsymbol{\beta}_c \sim \text{Normal}(0, \sigma_c^2 \textbf{I}) \\
& \qquad & \beta_{d,1} \sim \text{Normal}(0, 100^2) \\
& \qquad & \beta_{d,2} \sim \text{Normal}(0, 100^2) \\
& \qquad & \tilde{\beta_0} \sim \text{Normal}(0, 100^2) \\
& \qquad & \tilde{\beta_s} \sim \text{Normal}(0, 100^2) \\
& \qquad & \sigma_Q \sim \text{Cauchy}(0, 5) \\
& \qquad & \sigma_{\beta_0} \sim \text{Cauchy}(0, 5) \\
& \qquad & \sigma_{\beta_s} \sim \text{Cauchy}(0, 5)
\end{align}

@@

**Stan code for model**
```{r survival_stan, echo=TRUE, eval=FALSE}
data{
  // All Data
  int<lower=0> N; // observations
  int<lower=0> Yrs; // years
  int<lower=0> yid[N]; // year id
  int<lower=0> Covs; // climate covariates
  int<lower=0> G; // groups
  int<lower=0> gid[N]; // group id
  int<lower=0,upper=1> Y[N]; // observation vector
  matrix[N,Covs] C; // climate matrix
  vector[N] X; // size vector
  matrix[N,2] W; // crowding matrix
  real beta_tau; // prior sdev for climate effects
}
parameters{
  real a_mu;
  vector[Yrs] a;
  real b1_mu;
  vector[Yrs] b1;
  vector[Covs] b2;
  vector[2] w;
  vector[G] gint;
  real<lower=0> sig_a;
  real<lower=0> sig_b1;
  real<lower=0> sig_G;
}
transformed parameters{
  real mu[N];
  vector[N] climEff;
  vector[N] crowdEff;
  climEff <- C*b2;
  crowdEff <- W*w;
  for(n in 1:N){
    mu[n] <- inv_logit(a[yid[n]] + gint[gid[n]] + 
                         b1[yid[n]]*X[n] + crowdEff[n] + climEff[n]);
  }
}
model{
  // Priors
  a_mu ~ normal(0,100);
  w ~ normal(0,100);
  b1_mu ~ normal(0,100);
  sig_a ~ cauchy(0,5);
  sig_b1 ~ cauchy(0,5);
  sig_G ~ cauchy(0,5);
  gint ~ normal(0, sig_G);
  b2 ~ normal(0, beta_tau);
  a ~ normal(a_mu, sig_a);
  b1 ~ normal(b1_mu, sig_b1);

@@

**Data, process, and parameter models.**
\begin{align}
&\mathtt{data} \qquad &y_{i,q,t+1}^G \sim \text{Normal}(\mu_{i,q,t+1}, \sigma^2_{x_{i,q,t+1}}\textbf{I}) \\
&\mathtt{process} \qquad &\mu_{i,q,t+1} = \beta_{0,t} + \beta_{s,t}x_{i,q,t} + \beta_Q + \textbf{z}'_t \boldsymbol{\beta}_c + \beta_{d,1} w_{i,q,t} + \beta_{d,2} (x_{i,q,t}w_{i,q,t}) \\
& \qquad & \sigma^2_{iQ,t} = ae^{b\mu_{iQ,t+1}} \\
&\mathtt{parameters} \qquad &\beta_{0,t} \sim \text{Normal}(\tilde{\beta_0}, \sigma_{\beta_0}^2 \textbf{I})\\
& \qquad & \beta_{s,t} \sim \text{Normal}(\tilde{\beta_s}, \sigma_{\beta_s}^2 \textbf{I}) \\
& \qquad & \boldsymbol{\beta_{Q}} \sim \text{Normal}(0, \sigma_Q^2 \textbf{I}) \\
& \qquad & \boldsymbol{\beta}_c \sim \text{Normal}(0, \sigma_c^2 \textbf{I}) \\
& \qquad & \beta_{d,1} \sim \text{Normal}(0, 10^2) \\
& \qquad & \beta_{d,2} \sim \text{Normal}(0, 10^2) \\
& \qquad & \tilde{\beta_0} \sim \text{Normal}(0, 100^2) \\
& \qquad & \tilde{\beta_s} \sim \text{Normal}(0, 100^2) \\
& \qquad & \sigma_Q \sim \text{Cauchy}(0, 2) \\
& \qquad & \sigma_{\beta_0} \sim \text{Cauchy}(0, 2) \\
& \qquad & \sigma_{\beta_s} \sim \text{Cauchy}(0, 2) \\
& \qquad & a \sim \text{Normal}(0, 100^2) \\
& \qquad & b \sim \text{Normal}(0, 100^2)
\end{align}

@@

**Stan code for model**
```{r grow_stan, echo=TRUE, eval=FALSE}
data{
  int<lower=0> N; // observations
  int<lower=0> Yrs; // years
  int<lower=0> yid[N]; // year id
  int<lower=0> Covs; // climate covariates
  real<lower=0> tau_beta; // prior sdev for climate effects
  int<lower=0> G; // groups
  int<lower=0> gid[N]; // group id
  vector[N] Y; // observation vector
  matrix[N,Covs] C; // climate matrix
  vector[N] X; // size vector
  matrix[N,2] W; // crowding matrix
}
parameters{
  real a_mu;
  vector[Yrs] a;
  real b1_mu;
  vector[Yrs] b1;
  vector[Covs] b2;
  vector[2] w;
  real gint[G];
  real tau;
  real tauSize;
  real<lower=0> sig_a;
  real<lower=0> sig_b1;
  real<lower=0> sig_G;
}
transformed parameters{
  vector[N] mu;
  real<lower=0> sigma[N];
  vector[N] climEff;
  vector[N] crowdEff;
  climEff <- C*b2;
  crowdEff <- W*w;
  for(n in 1:N){
    mu[n] <- a[yid[n]] + gint[gid[n]] + b1[yid[n]]*X[n] + crowdEff[n] + climEff[n];
    sigma[n] <- sqrt((fmax(tau*exp(tauSize*mu[n]), 0.0000001)));  
  }
}
model{
  // Priors
  a_mu ~ normal(0,100);
  w ~ normal(0,10);
  b1_mu ~ normal(0,100);
  tau ~ normal(0,100);
  tauSize ~ normal(0,100);
  sig_a ~ cauchy(0,2);
  sig_b1 ~ cauchy(0,2);
  sig_G ~ cauchy(0,2);
  b2 ~ normal(0, tau_beta);
  for(g in 1:G)
    gint[g] ~ normal(0, sig_G);
  for(y in 1:Yrs){
    a[y] ~ normal(a_mu, sig_a);
    b1[y] ~ normal(b1_mu, sig_b1);
  }

@@

**Data, process, and parameter models**
\begin{align}
&\mathtt{data} \qquad &y^{R}_{q,t+1} \sim \text{NegBin}(\lambda_{q,t+1},\phi) \\
&\mathtt{process} \qquad &\lambda_{q,t+1} = \tilde{c}_{q,t} \, \text{exp}\left(\beta_{0,t} + \beta_Q + \textbf{z}'_t \boldsymbol{\beta}_c + \beta_{d}\sqrt{\tilde{c}_{q,t}}\right) \\
&\mathtt{parameters} \qquad &\beta_{0,t} \sim \text{Normal}(\tilde{\beta_0}, \sigma_{\beta_0}^2 \textbf{I})\\
& \qquad & \beta_{Q} \sim \text{Normal}(0,\sigma_Q^2 \textbf{I}) \\
& \qquad & \boldsymbol{\beta}_c \sim \text{Normal}(0, \sigma_c^2 \textbf{I}) \\
& \qquad & \beta_{d} \sim \text{Uniform}(-100, 100) \\
& \qquad & \tilde{\beta_0} \sim \text{Normal}(0, 1000^2) \\
& \qquad & \sigma_Q \sim \text{Cauchy}(0, 5) \\
& \qquad & \sigma_{\beta_0} \sim \text{Cauchy}(0, 5) \\
& \qquad & \phi \sim \text{Uniform}(0, 100) \\
& \qquad & u \sim \text{Uniform}(0, 1)
\end{align}

@@

**Stan code for model**
```{r rec_stan, echo=TRUE, eval=FALSE}
data{
  int<lower=0> N; // observations
  int<lower=0> Yrs; // years
  int<lower=0> yid[N]; // year id
  int<lower=0> Covs; // climate covariates
  int<lower=0> G; // groups
  int<lower=0> gid[N]; // group id
  int<lower=0> Y[N]; // observation vector
  matrix[N,Covs] C; // climate matrix
  vector[N] parents1; // crowding vector
  vector[N] parents2; // crowding vector
  real<lower=0> tau; // prior variance
}
parameters{
  real a_mu;
  vector[Yrs] a;
  vector[Covs] b2;
  real dd;
  real gint[G];
  real<lower=0> sig_a;
  real<lower=0> theta;
  real<lower=0> sig_G;
  real<lower=0, upper=1> u;
}
transformed parameters{
real mu[N];
vector[N] climEff;
vector[N] trueP1;
vector[N] trueP2;
vector[N] lambda;
vector[N] q;
climEff <- C*b2;
  for(n in 1:N){
    trueP1[n] <- parents1[n]*u + parents2[n]*(1-u);
    trueP2[n] <- sqrt(trueP1[n]);
    mu[n] <- exp(a[yid[n]] + gint[gid[n]] + dd*trueP2[n] + climEff[n]);
    lambda[n] <- trueP1[n]*mu[n];
    q[n] <- lambda[n]*theta;
  }
}
model{
// Priors
  u ~ uniform(0,1);
  theta ~ uniform(0,100);
  a_mu ~ normal(0,1000);
  dd ~ uniform(-100,100);
  sig_a ~ cauchy(0,5);
  sig_G ~ cauchy(0,5);
  for(g in 1:G)
   gint[g] ~ normal(0, sig_G);
  for(y in 1:Yrs){
    a[y] ~ normal(a_mu, sig_a);
  }
  for(j in 1:Covs)
    b2[j] ~ normal(0, tau);

@@

**Data, process, and parameter models.**
\begin{align}
&\mathtt{data} \qquad &y_{q,t+1}^{P} \sim \text{LogNormal}(\mu_{q,t+1}, \sigma^2)_{0}^{1} \\
&\mathtt{process} \qquad &\mu_{q,t+1} = \beta_{0,t} + \beta_{s,t}x_{q,t} + \beta_Q + \textbf{z}'_t \boldsymbol{\beta}_c \\
&\mathtt{parameters} \qquad &\beta_{0,t} \sim \text{Normal}(\tilde{\beta_0}, \sigma_{\beta_0}^2)\\
& \qquad & \beta_{s,t} \sim \text{Normal}(\tilde{\beta_s}, \sigma_{\beta_s}^2 \textbf{I}) \\
& \qquad & \beta_{Q} \sim \text{Normal}(0,\sigma_Q^2 \textbf{I}) \\
& \qquad & \boldsymbol{\beta}_c \sim \text{Normal}(0, \sigma_c^2 \textbf{I}) \\
& \qquad & \tilde{\beta_0} \sim \text{Normal}(0, 1000^2) \\
& \qquad & \tilde{\beta_s} \sim \text{Normal}(0, 1000^2) \\
& \qquad & \sigma_Q \sim \text{Cauchy}(0, 5) \\
& \qquad & \sigma_{\beta_0} \sim \text{Cauchy}(0, 5) \\
& \qquad & \sigma_{\beta_s} \sim \text{Cauchy}(0, 5) \\
& \qquad & \sigma \sim \text{Cauchy}(0, 5)
\end{align}

@@

**Stan code for model**
```{r qbm_stan, echo=TRUE, eval=FALSE}
data{
  int<lower=0> N; // observations
  int<lower=0> Yrs; // years
  int<lower=0> yid[N]; // year id
  int<lower=0> Covs; // climate covariates
  int<lower=0> G; // groups
  int<lower=0> gid[N]; // group id
  real<lower=0,upper=1> Y[N]; // observation vector
  real<lower=0> sd_clim; // prior sd on climate effects
  matrix[N,Covs] C; // climate matrix
  vector[N] X; // size vector
}
parameters{
  real a_mu;
  vector[Yrs] a;
  real b1_mu;
  vector[Yrs] b1;
  vector[Covs] b2;
  vector[G] gint;
  real<lower=0> sig_a;
  real<lower=0> sig_b1;
  real<lower=0> sig_G;
  real<lower=0> tau;
}
transformed parameters{
  real mu[N];
  vector[N] climEff;
  climEff <- C*b2;
  for(n in 1:N)
    mu[n] <- a[yid[n]] + gint[gid[n]] + b1[yid[n]]*X[n] + climEff[n];
}
model{
  // Priors
  a_mu ~ normal(0,1000);
  b1_mu ~ normal(0,1000);
  sig_a ~ cauchy(0,5);
  sig_b1 ~ cauchy(0,5);
  sig_G ~ cauchy(0,5);
  gint ~ normal(0, sig_G);
  b2 ~ normal(0,sd_clim);
  a ~ normal(a_mu, sig_a);
  b1 ~ normal(b1_mu, sig_b1);
  tau ~ cauchy(0,5);

>>>>

\begin{align}
&\mathtt{data} \qquad &y_{i,q,t}^{S} \sim \text{Bernoulli}(s_{i,q,t}) \\
&\mathtt{process} \qquad &\text{logit}(s_{i,q,t}) = \beta_{0,t} + \beta_{s,t}x_{i,q,t} + \beta_Q + \textbf{z}'_t \boldsymbol{\beta}_c + \beta_{d,1} w_{i,t} + \beta_{d,2} (x_{i,q,t}w_{i,q,t}) \\
&\mathtt{parameters} \qquad &\beta_{0,t} \sim \text{Normal}(\tilde{\beta_0}, \sigma_{\beta_0}^2 \textbf{I})\\
& \qquad & \beta_{s,t} \sim \text{Normal}(\tilde{\beta_s}, \sigma_{\beta_s}^2 \textbf{I}) \\
& \qquad & \boldsymbol{\beta_{Q}} \sim \text{Normal}(0, \sigma_Q^2 \textbf{I}) \\
& \qquad & \boldsymbol{\beta}_c \sim \text{Normal}(0, \sigma_c^2 \textbf{I}) \\
& \qquad & \beta_{d,1} \sim \text{Normal}(0, 100) \\
& \qquad & \beta_{d,2} \sim \text{Normal}(0, 100) \\
& \qquad & \tilde{\beta_0} \sim \text{Normal}(0, 100) \\
& \qquad & \tilde{\beta_s} \sim \text{Normal}(0, 100) \\
& \qquad & \sigma_Q \sim \text{Cauchy}(0, 5) \\
& \qquad & \sigma_{\beta_0} \sim \text{Cauchy}(0, 5) \\
& \qquad & \sigma_{\beta_s} \sim \text{Cauchy}(0, 5)
\end{align}

@@

**Stan code for model**
```{r survival_stan, echo=FALSE, eval=FALSE}
data{
  // All Data
  int<lower=0> N; // observations
  int<lower=0> Yrs; // years
  int<lower=0> yid[N]; // year id
  int<lower=0> Covs; // climate covariates
  int<lower=0> G; // groups
  int<lower=0> gid[N]; // group id
  int<lower=0,upper=1> Y[N]; // observation vector
  matrix[N,Covs] C; // climate matrix
  vector[N] X; // size vector
  matrix[N,2] W; // crowding matrix
  real beta_tau; // prior sdev for climate effects
}
parameters{
  real a_mu;
  vector[Yrs] a;
  real b1_mu;
  vector[Yrs] b1;
  vector[Covs] b2;
  vector[2] w;
  vector[G] gint;
  real<lower=0> sig_a;
  real<lower=0> sig_b1;
  real<lower=0> sig_G;
}
transformed parameters{
  real mu[N];
  vector[N] climEff;
  vector[N] crowdEff;
  climEff <- C*b2;
  crowdEff <- W*w;
  for(n in 1:N){
    mu[n] <- inv_logit(a[yid[n]] + gint[gid[n]] + 
                         b1[yid[n]]*X[n] + crowdEff[n] + climEff[n]);
  }
}
model{
  // Priors
  a_mu ~ normal(0,100);
  w ~ normal(0,100);
  b1_mu ~ normal(0,100);
  sig_a ~ cauchy(0,5);
  sig_b1 ~ cauchy(0,5);
  sig_G ~ cauchy(0,5);
  gint ~ normal(0, sig_G);
  b2 ~ normal(0, beta_tau);
  a ~ normal(a_mu, sig_a);
  b1 ~ normal(b1_mu, sig_b1);

@@

**Data, process, and parameter models.**
\begin{align}
&\mathtt{data} \qquad &y_{i,q,t+1}^G \sim \text{Normal}(\mu_{i,q,t+1}, \sigma^2_{x_{i,q,t+1}}\textbf{I}) \\
&\mathtt{process} \qquad &\mu_{i,q,t+1} = \beta_{0,t} + \beta_{s,t}x_{i,q,t} + \beta_Q + \textbf{z}'_t \boldsymbol{\beta}_c + \beta_{d,1} w_{i,q,t} + \beta_{d,2} (x_{i,q,t}w_{i,q,t}) \\
& \qquad & \sigma^2_{iQ,t} = ae^{b\mu_{iQ,t+1}} \\
&\mathtt{parameters} \qquad &\beta_{0,t} \sim \text{Normal}(\tilde{\beta_0}, \sigma_{\beta_0}^2 \textbf{I})\\
& \qquad & \beta_{s,t} \sim \text{Normal}(\tilde{\beta_s}, \sigma_{\beta_s}^2 \textbf{I}) \\
& \qquad & \boldsymbol{\beta_{Q}} \sim \text{Normal}(0, \sigma_Q^2 \textbf{I}) \\
& \qquad & \boldsymbol{\beta}_c \sim \text{Normal}(0, \sigma_c^2 \textbf{I}) \\
& \qquad & \beta_{d,1} \sim \text{Normal}(0, 100) \\
& \qquad & \beta_{d,2} \sim \text{Normal}(0, 100) \\
& \qquad & \tilde{\beta_0} \sim \text{Normal}(0, 100) \\
& \qquad & \tilde{\beta_s} \sim \text{Normal}(0, 100) \\
& \qquad & \sigma_Q \sim \text{Cauchy}(0, 2) \\
& \qquad & \sigma_{\beta_0} \sim \text{Cauchy}(0, 5) \\
& \qquad & \sigma_{\beta_s} \sim \text{Cauchy}(0, 5) \\
& \qquad & a \sim \text{Normal}(0, 100) \\
& \qquad & b \sim \text{Normal}(0, 100)
\end{align}

@@

**Stan code for model**
```{r grow_stan, echo=FALSE, eval=FALSE}
data{
  int<lower=0> N; // observations
  int<lower=0> Yrs; // years
  int<lower=0> yid[N]; // year id
  int<lower=0> Covs; // climate covariates
  real<lower=0> tau_beta; // prior sdev for climate effects
  int<lower=0> G; // groups
  int<lower=0> gid[N]; // group id
  vector[N] Y; // observation vector
  matrix[N,Covs] C; // climate matrix
  vector[N] X; // size vector
  matrix[N,2] W; // crowding matrix
}
parameters{
  real a_mu;
  vector[Yrs] a;
  real b1_mu;
  vector[Yrs] b1;
  vector[Covs] b2;
  vector[2] w;
  real gint[G];
  real tau;
  real tauSize;
  real<lower=0> sig_a;
  real<lower=0> sig_b1;
  real<lower=0> sig_G;
}
transformed parameters{
  vector[N] mu;
  real<lower=0> sigma[N];
  vector[N] climEff;
  vector[N] crowdEff;
  climEff <- C*b2;
  crowdEff <- W*w;
  for(n in 1:N){
    mu[n] <- a[yid[n]] + gint[gid[n]] + b1[yid[n]]*X[n] + crowdEff[n] + climEff[n];
    sigma[n] <- sqrt((fmax(tau*exp(tauSize*mu[n]), 0.0000001)));  
  }
}
model{
  // Priors
  a_mu ~ normal(0,100);
  w ~ normal(0,10);
  b1_mu ~ normal(0,100);
  tau ~ normal(0,100);
  tauSize ~ normal(0,100);
  sig_a ~ cauchy(0,2);
  sig_b1 ~ cauchy(0,2);
  sig_G ~ cauchy(0,2);
  b2 ~ normal(0, tau_beta);
  for(g in 1:G)
    gint[g] ~ normal(0, sig_G);
  for(y in 1:Yrs){
    a[y] ~ normal(a_mu, sig_a);
    b1[y] ~ normal(b1_mu, sig_b1);
  }

@@

**Data, process, and parameter models**
\begin{align}
&\mathtt{data} \qquad &y^{R}_{q,t+1} \sim \text{NegBin}(\lambda_{q,t+1},\phi) \\
&\mathtt{process} \qquad &\lambda_{q,t+1} = \tilde{c}_{q,t} \, \text{exp}\left(\beta_{0,t} + \beta_Q + \textbf{z}'_t \boldsymbol{\beta}_c + \beta_{d}\sqrt{\tilde{c}_{q,t}}\right) \\
&\mathtt{parameters} \qquad &\beta_{0,t} \sim \text{Normal}(\tilde{\beta_0}, \sigma_{\beta_0}^2 \textbf{I})\\
& \qquad & \beta_{Q} \sim \text{Normal}(0,\sigma_Q^2 \textbf{I}) \\
& \qquad & \boldsymbol{\beta}_c \sim \text{Normal}(0, \sigma_c^2 \textbf{I}) \\
& \qquad & \beta_{d} \sim \text{Uniform}(-10, 10) \\
& \qquad & \tilde{\beta_0} \sim \text{Normal}(0, 100) \\
& \qquad & \sigma_Q \sim \text{Cauchy}(0, 5) \\
& \qquad & \sigma_{\beta_0} \sim \text{Cauchy}(0, 5) \\
& \qquad & \phi \sim \text{Uniform}(0, 10) \\
& \qquad & u \sim \text{Uniform}(0, 1)
\end{align}

@@

**Stan code for model**
```{r rec_stan, echo=FALSE, eval=FALSE}
data{
  int<lower=0> N; // observations
  int<lower=0> Yrs; // years
  int<lower=0> yid[N]; // year id
  int<lower=0> Covs; // climate covariates
  int<lower=0> G; // groups
  int<lower=0> gid[N]; // group id
  int<lower=0> Y[N]; // observation vector
  matrix[N,Covs] C; // climate matrix
  vector[N] parents1; // crowding vector
  vector[N] parents2; // crowding vector
  real<lower=0> tau; // prior variance
}
parameters{
  real a_mu;
  vector[Yrs] a;
  vector[Covs] b2;
  real dd;
  real gint[G];
  real<lower=0> sig_a;
  real<lower=0> theta;
  real<lower=0> sig_G;
  real<lower=0, upper=1> u;
}
transformed parameters{
real mu[N];
vector[N] climEff;
vector[N] trueP1;
vector[N] trueP2;
vector[N] lambda;
vector[N] q;
climEff <- C*b2;
  for(n in 1:N){
    trueP1[n] <- parents1[n]*u + parents2[n]*(1-u);
    trueP2[n] <- sqrt(trueP1[n]);
    mu[n] <- exp(a[yid[n]] + gint[gid[n]] + dd*trueP2[n] + climEff[n]);
    lambda[n] <- trueP1[n]*mu[n];
    q[n] <- lambda[n]*theta;
  }
}
model{
// Priors
  u ~ uniform(0,1);
  theta ~ uniform(0,100);
  a_mu ~ normal(0,1000);
  dd ~ uniform(-100,100);
  sig_a ~ cauchy(0,5);
  sig_G ~ cauchy(0,5);
  for(g in 1:G)
   gint[g] ~ normal(0, sig_G);
  for(y in 1:Yrs){
    a[y] ~ normal(a_mu, sig_a);
  }
  for(j in 1:Covs)
    b2[j] ~ normal(0, tau);

@@

**Data, process, and parameter models.**
\begin{align}
&\mathtt{data} \qquad &y_{q,t+1}^{P} \sim \text{LogNormal}(\mu_{q,t+1}, \sigma^2)_{0}^{1} \\
&\mathtt{process} \qquad &\mu_{q,t+1} = \beta_{0,t} + \beta_{s,t}x_{q,t} + \beta_Q + \textbf{z}'_t \boldsymbol{\beta}_c \\
&\mathtt{parameters} \qquad &\beta_{0,t} \sim \text{Normal}(\tilde{\beta_0}, \sigma_{\beta_0}^2)\\
& \qquad & \beta_{s,t} \sim \text{Normal}(\tilde{\beta_s}, \sigma_{\beta_s}^2 \textbf{I}) \\
& \qquad & \beta_{Q} \sim \text{Normal}(0,\sigma_Q^2 \textbf{I}) \\
& \qquad & \boldsymbol{\beta}_c \sim \text{Normal}(0, \sigma_c^2 \textbf{I}) \\
& \qquad & \tilde{\beta_0} \sim \text{Normal}(0, 100) \\
& \qquad & \tilde{\beta_s} \sim \text{Normal}(0, 100) \\
& \qquad & \sigma_Q \sim \text{Cauchy}(0, 5) \\
& \qquad & \sigma_{\beta_0} \sim \text{Cauchy}(0, 5) \\
& \qquad & \sigma_{\beta_s} \sim \text{Cauchy}(0, 5) \\
& \qquad & \sigma^2 \sim \text{Inverse Gamma}(0.001, 0.001)
\end{align}

@@

**Stan code for model**
```{r qbm_stan, echo=FALSE, eval=FALSE}
data{
  int<lower=0> N; // observations
  int<lower=0> Yrs; // years
  int<lower=0> yid[N]; // year id
  int<lower=0> Covs; // climate covariates
  int<lower=0> G; // groups
  int<lower=0> gid[N]; // group id
  real<lower=0,upper=1> Y[N]; // observation vector
  real<lower=0> sd_clim; // prior sd on climate effects
  matrix[N,Covs] C; // climate matrix
  vector[N] X; // size vector
}
parameters{
  real a_mu;
  vector[Yrs] a;
  real b1_mu;
  vector[Yrs] b1;
  vector[Covs] b2;
  vector[G] gint;
  real<lower=0> sig_a;
  real<lower=0> sig_b1;
  real<lower=0> sig_G;
  real<lower=0> tau;
}
transformed parameters{
  real mu[N];
  vector[N] climEff;
  climEff <- C*b2;
  for(n in 1:N)
    mu[n] <- a[yid[n]] + gint[gid[n]] + b1[yid[n]]*X[n] + climEff[n];
}
model{
  // Priors
  a_mu ~ normal(0,1000);
  b1_mu ~ normal(0,1000);
  sig_a ~ cauchy(0,5);
  sig_b1 ~ cauchy(0,5);
  sig_G ~ cauchy(0,5);
  gint ~ normal(0, sig_G);
  b2 ~ normal(0,sd_clim);
  a ~ normal(a_mu, sig_a);
  b1 ~ normal(b1_mu, sig_b1);
  tau ~ cauchy(0,5);
